---
title: "quantitative metabarcoding figures for mock communities"
author: "Kimberly Ledger"
date: "2023-10-06"
output: html_document
---

load libraries
```{r}
library(tidyverse)
library(data.table)
library(gridExtra)
library(ggsci)
```

# Read in the posteriors for mock communities
```{r}
load("/home/kimberly.ledger/gadid_aquaria/gadid_mb_mockcommunity/quant_outputs/raw_20231016_even.Rdata")
raw <- Output

load("/home/kimberly.ledger/gadid_aquaria/gadid_mb_mockcommunity/quant_outputs/mock_20231016_even.Rdata")
mock1 <- Output
```


```{r}
#########################################################
# Raw estimates from Reads
#########################################################
# summarize raw estimates from reads for each species.
raw.raw <- raw$env %>% 
  group_by(community,Cycles,tech_rep) %>%
  mutate(sum.ng = sum(start_conc_ng),
         true.prop = start_conc_ng / sum.ng) %>%
  ungroup() %>%
  group_by(Species,community,Cycles,true.prop) %>%
  group_by(Species,community,Cycles) %>%
  summarize(simple.Mean=mean(propReads),
            simple.N = length(tech_rep)) %>%
  replace_na(list(raw.Mean=0,raw.SD=0,raw.SE=0))

# extract predicted proportions from the posterior
COM <- data.frame(community = levels(raw$env$community %>% as.factor()))
COM$comm_idx <- 1:nrow(COM)
SP  <- raw$env %>% distinct(Species,sp_idx) %>% as.data.frame()

# These are the predicted intercepts for the posteriors
beta_posterior <- raw$stanMod_summary[["int_samp_small"]][, c(1,4:8)]
colnames(beta_posterior) <- paste0("raw.",substr(colnames(beta_posterior),1,nchar(colnames(beta_posterior))-1))
colnames(beta_posterior)[1] <- "raw.mean"
beta_posterior <- as.data.frame(beta_posterior)

raw.post <-expand.grid(comm_idx = COM$comm_idx,sp_idx =SP$sp_idx) %>% 
  arrange(comm_idx,sp_idx) %>% 
  left_join(.,COM) %>% 
  left_join(.,SP) %>% 
  bind_cols(.,beta_posterior)

# Combine the raw estimates and posterior estimates
raw.all <- full_join(raw.raw,raw.post)
```
```{r}
#########################################################
# Mock1
#########################################################
# summarize raw estimates from reads for each species.
mock1.raw <- mock1$env %>% group_by(community,Cycles,tech_rep) %>%
      mutate(sum.ng = sum(start_conc_ng),
             true.prop = start_conc_ng / sum.ng) %>%
      ungroup() %>%
      group_by(Species,community,Cycles,true.prop) %>%
  summarize(simple.Mean=mean(propReads),
            simple.N = length(tech_rep)) %>%
  replace_na(list(raw.Mean=0,raw.SD=0,raw.SE=0))

# extract predicted proportions from the posterior
COM <- data.frame(community = levels(mock1$env$community %>% as.factor()))
COM$comm_idx <- 1:nrow(COM)
SP  <- mock1$env %>% distinct(Species,sp_idx) %>% as.data.frame()

# These are the predicted intercepts for the posteriors
beta_posterior <- mock1$stanMod_summary[["int_samp_small"]][, c(1,4:8)]
colnames(beta_posterior) <- paste0("mock1.",substr(colnames(beta_posterior),1,nchar(colnames(beta_posterior))-1))
colnames(beta_posterior)[1] <- "mock1.mean"
beta_posterior <- as.data.frame(beta_posterior)

mock1.post <-expand.grid(comm_idx = COM$comm_idx,sp_idx =SP$sp_idx) %>% 
    arrange(comm_idx,sp_idx) %>% 
    left_join(.,COM) %>% 
    left_join(.,SP) %>% 
    bind_cols(.,beta_posterior)

# Combine the raw estimates and posterior estimates
mock1.all <- full_join(mock1.raw,mock1.post)
```

```{r}
# Combine mock results with raw reads.
result.dat <- left_join(mock1.all,raw.all)
```

plot how each of the mock communities set as unknown performed 
```{r}
# pull out just M1 for plotting.
spread=0.15
M1.dat <- result.dat %>% 
  filter(true.prop > 0,community=="middle")
M1.dat <- bind_cols(M1.dat, data.frame(offset= seq(-spread,spread,length.out=nrow(M1.dat))))
# pull out just N1 for plotting.
N1.dat <- result.dat %>% 
  filter(true.prop > 0,community=="north1")
N1.dat <- bind_cols(N1.dat, data.frame(offset= seq(-spread,spread,length.out=nrow(N1.dat))))
# pull out just S1 for plotting.
S1.dat <- result.dat %>% 
  filter(true.prop > 0,community=="south1")
S1.dat <- bind_cols(S1.dat,data.frame(offset= seq(-spread,spread,length.out=nrow(S1.dat))))
# pull out just N2 for plotting.
N2.dat <- result.dat %>% 
  filter(true.prop > 0,community=="north2")
N2.dat <- bind_cols(N2.dat, data.frame(offset= seq(-spread,spread,length.out=nrow(N2.dat))))
# pull out just S2 for plotting.
S2.dat <- result.dat %>% 
  filter(true.prop > 0,community=="south2")
S2.dat <- bind_cols(S2.dat,data.frame(offset= seq(-spread,spread,length.out=nrow(S2.dat))))

# Make plots
BREAKS <- c(0.0,0.05,0.10,0.25,0.50)
x.labs <- c("None","Mock")                          ## not sure what this is designating later on 
x.at   <- c(1,2)

skew_plot <- function(dat,
                      BREAKS=BREAKS,x.labs=x.labs,x.at=x.at){
  
  #shape.val = c(21,22)
  #col.val = pal_jco(palette = c("default"), alpha = 1)(10)[c(1,4)]
  
  skew.plot <-  ggplot(dat) +
    scale_color_gradientn(colours = rainbow(4)) +
    geom_errorbar(aes(x=1+offset,
                      ymin=raw.2.5,   
                      ymax=raw.97.5,color=true.prop),width=0,alpha=0.75)   +
    geom_point(aes(x=1+offset,y=raw.mean, #shape=true.prop,
                   fill=true.prop,color=true.prop),size=2) +
    # mock with mock communities at multiple PCR
     geom_errorbar(aes(x=2+offset,
                       ymin= mock1.2.5, 
                       ymax= mock1.97.5, color = true.prop),width=0,alpha=0.5)   +
     geom_point(aes(x=2+offset,mock1.mean, #shape=true.prop,
                    fill=true.prop,color=true.prop), size=2) +
    #scale_shape_manual(values =c(21,22)) +
    #scale_fill_manual(values= col.val, "True value") +
    #scale_color_manual(values= col.val,"True value") +
    scale_y_continuous("Proportion",
                       trans="sqrt",
                       # trans="log",
                       breaks = BREAKS,expand=c(0,NA),limits = c(0,NA)) +
    geom_hline(aes(yintercept = true.prop,color=true.prop),linetype="dashed") +
    #geom_point(aes(x=0.70,y=true.prop,shape=true.prop,fill=true.prop),size=3) +
    scale_x_continuous(name=NULL,breaks=x.at,labels = x.labs) +
    theme_classic() +
    theme(legend.position = "none")
  
  return(skew.plot)
}


M1_plot <- skew_plot(dat=M1.dat,
                          BREAKS=BREAKS,x.labs=x.labs,x.at=x.at) 
M1_plot                

N1_plot <- skew_plot(dat=N1.dat,
                          BREAKS=BREAKS,x.labs=x.labs,x.at=x.at) 
S1_plot <- skew_plot(dat=S1.dat,
                          BREAKS=BREAKS,x.labs=x.labs,x.at=x.at) 
N2_plot <- skew_plot(dat=N2.dat,
                          BREAKS=BREAKS,x.labs=x.labs,x.at=x.at) 

S2_plot <- skew_plot(dat=S2.dat,
                          BREAKS=BREAKS,x.labs=x.labs,x.at=x.at) 

p3 <- grid.arrange(M1_plot + ggtitle(NULL,subtitle="Middle1"),
              N1_plot +ggtitle(NULL,subtitle="North1"),
              N2_plot +ggtitle(NULL,subtitle="North2"),
              S1_plot + ggtitle(NULL,subtitle="South1"),
              S2_plot + ggtitle(NULL,subtitle="South2"),
              ncol=5,nrow=1)
#ggsave(p3, filename = "/home/kimberly.ledger/gadid_metabarcoding/mock_community/mockcom_comparison/gadid_even_meanestimates.jpg", width = 9, height = 5, units = "in")
```


```{r}
# Make plots
BREAKS <- c(0.0,0.05,0.10,0.25,0.50)
x.labs <- c("Uncorrected","Corrected")                          
x.at   <- c(1,2)

## let me try to squish this into one plot 
all.dat <- M1.dat %>%
  bind_rows(N1.dat) %>%
  bind_rows(N2.dat) %>%
  bind_rows(S1.dat) %>%
  bind_rows(S2.dat) %>%
  mutate(Species = ifelse(Species == "zRefSpecies_Arctogadus glacialis", "Polar cod", Species)) %>%
  mutate(Species = ifelse(Species == "Boreogadus saida", "Arctic cod", Species)) %>%
  mutate(Species = ifelse(Species == "Eleginus gracilis", "Saffron cod", Species)) %>%
  mutate(Species = ifelse(Species == "Gadus chalcogrammus", "Walleye pollock", Species)) %>%
  mutate(Species = ifelse(Species == "Gadus macrocephalus", "Pacific cod", Species)) %>%
  mutate(Species = ifelse(Species == "Microgadus proximus", "Pacific tomcod", Species))

skew_plot <- function(dat,
                      BREAKS=BREAKS,x.labs=x.labs,x.at=x.at){
  
  #shape.val = c(21,22)
  #col.val = pal_jco(palette = c("default"), alpha = 1)(10)[c(1,4)]
  
  skew.plot <-  ggplot(dat) +
    scale_color_gradientn(colours = rainbow(8)[c(1,2,3,6)]) +
    geom_errorbar(aes(x=1+offset,
                      ymin=raw.2.5,   
                      ymax=raw.97.5,color=true.prop),width=0,alpha=0.75)   +
    geom_point(aes(x=1+offset,y=raw.mean, shape=Species,
                   fill=true.prop,color=true.prop),size=2) +
    # mock with mock communities at multiple PCR
     geom_errorbar(aes(x=2+offset,
                       ymin= mock1.2.5, 
                       ymax= mock1.97.5, color = true.prop),width=0,alpha=0.5)   +
     geom_point(aes(x=2+offset,mock1.mean, shape=Species,
                    fill=true.prop,color=true.prop), size=2) +
    scale_shape_manual(values =c(16,15,17,1,8,18)) +
    #scale_fill_manual(values= col.val, "True value") +
    #scale_color_manual(values= col.val,"True value") +
    scale_y_continuous("Proportion",
                       trans="sqrt",
                       # trans="log",
                       breaks = BREAKS,expand=c(0,NA),limits = c(0,NA)) +
    geom_hline(aes(yintercept = true.prop,color=true.prop),linetype="dashed") +
    #geom_point(aes(x=0.70,y=true.prop,shape=true.prop,fill=true.prop),size=3) +
    scale_x_continuous(name=NULL,breaks=x.at,labels = x.labs) +
    theme_classic() +
    theme(legend.position = "right",
          legend.text = element_text(size = 8),
          legend.title = element_text(size = 9)) +
    guides(color = "none", fill = "none", shape = guide_legend(override.aes = list(size = 2)))
  
  return(skew.plot)
}

all_plot <- skew_plot(dat=all.dat,
                          BREAKS=BREAKS,x.labs=x.labs,x.at=x.at) 
all_plot

#ggsave(plot= all_plot, 
#      filename = ("~/gadid_aquaria/gadid_mb_mockcommunity/mock_adjustments.png"),
#      width=5,
#      height = 3,
#      dpi = 300,
#      units = c("in"))
```


let's visualize as scatter plots to match how we look at the tank experiment output
```{r}
my_colors <- c("#6a6599", "#79af97", "#df8f44", "#898E8C", "#b24746", "#00a1d5") 
#purple, green, orange, grey, red, blue

all.dat$Species <- as.factor(all.dat$Species)
levels(all.dat$Species)
```


#raw data 
```{r}
raw_plot <- all.dat %>%
  ggplot() + 
  geom_errorbar(aes(x=true.prop,ymin=raw.2.5,ymax=raw.97.5,color=Species),width=0, size = .5) +
  geom_point(aes(x=true.prop,y=raw.mean,color=Species), size = 2) +
  geom_smooth(aes(x=true.prop,y=raw.mean,color=Species), method = "lm", formula = y ~ x, se = FALSE) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black")  + 
  scale_color_manual(values = my_colors) + 
  theme_classic() + 
    labs(x = "true proportion",
       y = "uncorrected eDNA read proportion") +
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        legend.title = element_text(size = 12))

raw_plot
```

#mock-adjusted data 
```{r}
mock_plot <- all.dat %>%
  ggplot() + 
  geom_errorbar(aes(x=true.prop,ymin=mock1.2.5,ymax=mock1.97.5,color=Species),width=0, size = .5) +
  geom_point(aes(x=true.prop,y=mock1.mean,color=Species), size = 2) +
  geom_smooth(aes(x=true.prop,y=mock1.mean,color=Species), method = "lm", formula = y ~ x, se = FALSE) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black")  + 
  scale_color_manual(values = my_colors) + 
  theme_classic() + 
    labs(x = "true proportion",
       y = "corrected eDNA read proportion") +
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        legend.title = element_text(size = 12))
mock_plot
```

put together into one plot 
```{r, fig.width=10, fig.height=4}
library(patchwork)

combined_fig <- raw_plot + mock_plot + plot_layout(guides = 'collect') + plot_annotation(tag_levels = 'A') &
    theme(plot.tag = element_text(face = 'bold'))
combined_fig

#ggsave("raw_v_mock.png", plot = combined_fig, width = 10, height = 4, dpi = 300)
```

now let's calculate slopes/r2 for those models
```{r}
arctic <- all.dat %>%
  filter(Species == "Arctic cod")
saffron <- all.dat %>%
  filter(Species == "Saffron cod")
pollock <- all.dat %>%
  filter(Species == "Walleye pollock")
pacific <- all.dat %>%
  filter(Species == "Pacific cod")
tomcod <- all.dat %>%
  filter(Species == "Pacific tomcod")
polar <- all.dat %>%
  filter(Species == "Polar cod")

arctic_raw <- lm(raw.mean ~ true.prop, data = arctic)
summary(arctic_raw)
arctic_mock <- lm(mock1.mean ~ true.prop, data = arctic)
summary(arctic_mock)

saffron_raw <- lm(raw.mean ~ true.prop, data = saffron)
summary(saffron_raw)
saffron_mock <- lm(mock1.mean ~ true.prop, data = saffron)
summary(saffron_mock)

pollock_raw <- lm(raw.mean ~ true.prop, data = pollock)
summary(pollock_raw)
pollock_mock <- lm(mock1.mean ~ true.prop, data = pollock)
summary(pollock_mock)

pacific_raw <- lm(raw.mean ~ true.prop, data = pacific)
summary(pacific_raw)
pacific_mock <- lm(mock1.mean ~ true.prop, data = pacific)
summary(pacific_mock)

tomcod_raw <- lm(raw.mean ~ true.prop, data = tomcod)
summary(tomcod_raw)
tomcod_mock <- lm(mock1.mean ~ true.prop, data = tomcod)
summary(tomcod_mock)

polar_raw <- lm(raw.mean ~ true.prop, data = polar)
summary(polar_raw)
polar_mock <- lm(mock1.mean ~ true.prop, data = polar)
summary(polar_mock)
```


according to the even mock community, let me plot the proportional abundance pre-PCR vs post-PCR 
```{r}
fig_df <- all.dat %>%
  select(Species, community, true.prop, raw.mean, mock1.mean) %>%
  pivot_longer(cols = c(4:6), names_to = "method", values_to = "proportion") %>%
  unite("ID", Species, community, sep = "_", remove = F) %>%
  mutate(method = ifelse(method == "true.prop", "pre-PCR", method),
         method = ifelse(method == "mock1.mean", "corrected", method),
         method = ifelse(method == "raw.mean", "post-PCR", method)) 
  
fig_df$method <- factor(fig_df$method , levels = c("pre-PCR", "post-PCR", "corrected"))  

north2 <- fig_df %>%
  filter(community == "north2") %>%
  ggplot(aes(x = method, y = proportion, color = Species, group = ID)) + 
  geom_point() + 
  geom_line() +
  scale_color_manual(values = my_colors) + 
  theme_classic() + 
  labs(x = "",
       y = "proportional abundance (%)") +
  scale_x_discrete(expand = expansion(add = c(0.1, 0.1))) + 
  theme(axis.text = element_text(size = 10),
        axis.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        legend.title = element_text(size = 12),
        axis.text.x = element_text(angle = 90, hjust = 0.95)) +
  theme(legend.position = "none")

middle <- fig_df %>%
  filter(community == "middle") %>%
  ggplot(aes(x = method, y = proportion, color = Species, group = ID)) + 
  geom_point() + 
  geom_line() +
  scale_color_manual(values = my_colors) + 
  theme_classic() + 
  labs(x = "",
       y = "proportional abundance (%)") +
  scale_x_discrete(expand = expansion(add = c(0.1, 0.1))) + 
  theme(axis.text = element_text(size = 10),
        axis.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        legend.title = element_text(size = 12),
        axis.text.x = element_text(angle = 90, hjust = 0.95)) +
  theme(legend.position = "none")

north2_bar <- fig_df %>%
  filter(community == "north2") %>%
  ggplot(aes(x=method, y=proportion, fill=Species)) +
  geom_bar(stat = "identity") + 
  theme_bw() +
  scale_fill_manual(values = my_colors) + 
  #facet_grid(~community, scales = 'free') +
  labs(
    y = "proportional abundance (%)",
    x = "") + 
  theme(axis.text = element_text(size = 10),
        axis.title = element_text(size = 12),
    axis.text.x = element_text(angle = 90, hjust = 0.95),
    legend.text = element_text(size = 10),
    legend.key.size = unit(0.3, "cm"),
    legend.position = "right"
  )

middle_bar <- fig_df %>%
  filter(community == "middle") %>%
  ggplot(aes(x=method, y=proportion, fill=Species)) +
  geom_bar(stat = "identity") + 
  theme_bw() +
  scale_fill_manual(values = my_colors) + 
  #facet_grid(~community, scales = 'free') +
  labs(
    y = "proportional abundance (%)",
    x = "") + 
  theme(axis.text = element_text(size = 10),
        axis.title = element_text(size = 12),
    axis.text.x = element_text(angle = 90, hjust = 0.95),
    legend.text = element_text(size = 10),
    legend.key.size = unit(0.3, "cm"),
    legend.position = "right"
  )
```

put together into one plot 
```{r, fig.width=10, fig.height=4}
library(patchwork)

fig2 <- north2 + north2_bar + middle  + middle_bar + plot_layout(guides = 'collect') + plot_annotation(tag_levels = 'A') &
    theme(plot.tag = element_text(face = 'bold'))
fig2

ggsave("fig2.png", plot = fig2, width = 10, height = 8, dpi = 300)
```


additional plots

```{r}
my_colors <- c("#898E8C", "#6a6599", "#b24746", "#00a1d5", "#79af97", "#df8f44") 
#grey, purple, red, blue, green, orange
```


plot each of the unknown mock communities - true, mean raw, and mean mock-adjusted 
```{r}
temp1 <- result.dat %>%
  dplyr::select(Species, community, true.prop, mock1.mean, raw.mean)

#temp1 <- raw.post %>%
#  dplyr::select(Species, community, raw.mean)

temp2 <- temp1[,c(2:6)] %>%
  pivot_longer(cols = c(3:5), names_to = "method", values_to = "prop")

# get rid of reference species denotation
temp2 <- temp2 %>%
  mutate(SP= ifelse(grepl("zRefSpecies_",Species),
  substr(Species,13,nchar(Species)),
  as.character(Species)))

# rename "method"
temp3 <- temp2 %>%
  mutate(method = ifelse(method == "true.prop", "true proportion", method),
         method = ifelse(method == "mock1.mean", "mock adjusted", method),
         method = ifelse(method == "raw.mean", "raw proportion", method))

my_order <- c("true proportion", "raw proportion", "mock adjusted")
temp3$method <- factor(temp3$method, levels = my_order)

my_order_communities <- c("north1", "north2", "middle", "south1", "south2")
temp3$community <- factor(temp3$community, levels = my_order_communities)

plot_prop <- temp3 %>%
  ggplot(aes(x=method, y=prop, fill=SP)) +
  geom_bar(stat = "identity") + 
  theme_bw() +
  scale_fill_manual(values = my_colors) + 
  facet_grid(~community, scales = 'free') +
  labs(
    y = "proportion of sequencing reads",
    x = "method",
    title = "even as true mock") + 
  theme(
    axis.text.x = element_text(angle = 90, hjust = 0.95),
    legend.text = element_text(size = 8),
    legend.key.size = unit(0.3, "cm"),
    legend.position = "right",
    legend.title = element_blank()
  )

plot_prop
```

```{r}
ggsave("~/gadid_aquaria/gadid_mb_mockcommunity/evenmock_simple.png", plot = plot_prop, width = 6, height = 5, dpi = 300)
```


```{r}
###############################################################333
#### Pull out estimates of alpha, convert to CLR                        
###############################################################333


p_space_mock1 <- (exp(mock1$pars$alpha) / rowSums(exp(mock1$pars$alpha))) %>% as.data.frame()

clr_alpha_list_mock <- compositions::clr(p_space_mock1)

clr_alpha <- clr_alpha_list_mock %>% as.data.frame()

mock.sp   <- data.frame(Species=mock1.post %>% distinct(Species) %>% pull(Species))

colnames(clr_alpha) <- mock.sp$Species                       
clr_alpha_sum <- clr_alpha %>% 
                    pivot_longer( .,
                          cols = colnames(clr_alpha),
                          names_to="species",values_to="val") %>%
                    group_by(species) %>%
                    summarize(Mean = mean(val),
                        SD=sd(val),
                        q.025 = quantile(val,probs=0.025),
                        q.05 = quantile(val,probs=0.05),
                        q.25 = quantile(val,probs=0.25),
                        q.75 = quantile(val,probs=0.75),
                        q.95 = quantile(val,probs=0.95),
                        q.975 = quantile(val,probs=0.975))    

clr_alpha_sum <- clr_alpha_sum %>% arrange(Mean)

# get rid of reference species denotion
clr_alpha_sum <- clr_alpha_sum %>%
                      mutate(SP= ifelse(grepl("zRefSpecies_",species),
                                       substr(species,13,nchar(species)),
                                       as.character(species)))

p_clr_mock2 <-  ggplot(clr_alpha_sum) +
    geom_errorbarh(aes(xmin=q.25,xmax=q.75,y=SP),size=2,height=0) +
    geom_errorbarh(aes(xmin=q.025,xmax=q.975,y=SP),size=0.8,height=0) +
    geom_point(aes(x=Mean,y=SP,fill=SP,),size=3,shape=21) +
    geom_vline(xintercept=0,linetype="dashed") +
    #scale_fill_manual(values=clr_alpha_sum$manual.col %>% as.character()) +
    scale_x_continuous("Amplification Efficiency (CLR)") +
    scale_y_discrete(NULL) +
    theme_bw() +
    theme(legend.position = "none",
          axis.text.y = element_text(size=10))

p_clr_mock2

#ggsave(p_clr_mock2, filename = "/home/kimberly.ledger/gadid_metabarcoding/aquaria_DBO/quant_outputs/gadid_even_CLR.pdf", width = 5, height = 5, units = "in")
```

```{r}
clr_alpha_sum <- clr_alpha_sum %>%
  mutate(mock = "even")

#write.csv(clr_alpha_sum, "/home/kimberly.ledger/gadid_metabarcoding/aquaria_DBO/quant_outputs/gadid_even_CLR.csv", row.names = F)
```

