---
title: "tracking replicates/samples"
output: html_document
date: "2024-03-13"
---

load libraries
```{r}
library(tidyverse)
library(ggplot2)
select <- dplyr::select
rename <- dplyr::rename
```

read in starting replicate/sample metadata for the aquarium samples
```{r}
rep_metadata <- read.csv("gadid_mb_aquaria/gadid_aquariaDBO_metadata.csv") %>%
  filter(project == "aquaria") %>%
  select(!station_ID) %>%
  select(!location1:time_of_day)

#illumina output changed "_" to "-"
rep_metadata$Sample_ID <- gsub("_", "-", rep_metadata$Sample_ID) 
```

- replicates in MiSeq_run == A were decontamined in "1_20230918_ASV_decontamination_gadids.Rmd" 
- replicates in MiSeq_run == B were decontamined in "1_20230921_ASV_decontamination_gadids.Rmd" 

- replicates removed during decontamination step3 (b/c of low read numbers) are contained in "20230918_decontamination_step3.csv" and "20230921_decontamination_step3.csv"
```{r}
A_step3 <- read.csv("gadid_mb_aquaria/20230918_decontamination_step3.csv")
B_step3 <- read.csv("gadid_mb_aquaria/20230921_decontamination_step3.csv")

rep_metadata <- rep_metadata %>%
  mutate(low_reads = ifelse(Sample_ID %in% A_step3$Sample_ID, "yes", "no"),
         low_reads = ifelse(Sample_ID %in% B_step3$Sample_ID, "yes", low_reads))
```

- then 3 replicates/extractions were removed because only one PCR replicate of the extraction ID remained - see "3_combining_gadid_aquaria_runs.Rmd" 
```{r}
e02165-B
e02398-C
e02427-C
```

- flagged PCR replicates with group centroid outside 95% of the normal distribution of group distances
```{r}
pcr_outilers <- read.csv("gadid_mb_aquaria/pcr_dissimilarity_outliers.csv")

rep_metadata <- rep_metadata %>%
  unite(ID1, tank_ID, alt_ID, sep = "_", remove = FALSE) %>%
  unite(ID2, ID1, pcr_replicate, sep = "-", remove = FALSE) %>%
  mutate(pcr_replicate_outlier = ifelse(ID2 %in% pcr_outilers$discard_centroid, "yes", "no"),
         pcr_replicate_outlier = ifelse(low_reads == "yes", NA, pcr_replicate_outlier))
```

- tank dissimiarly - of mock1 mean estimates 
```{r}
tank_outliers <- read.csv("gadid_mb_aquaria/centroid_mock_tank_dissimilarity.csv")

rep_metadata <- rep_metadata %>%
  unite(ID3, tank_ID, extraction_ID, sep = "_", remove = FALSE) %>%
    mutate(tank_replicate_outlier = ifelse(ID3 %in% tank_outliers$x, "yes", "no"),
         tank_replicate_outlier = ifelse(low_reads == "yes", NA, tank_replicate_outlier))
```


```{r}
dna_conc_sample_type <- rep_metadata %>%
  filter(sample_type != "positive") %>%
   filter(sample_type != "PCR_blank") %>%
  ggplot(aes(x = sample_type, y = DNA_conc, fill = low_reads)) +
  geom_boxplot() + 
  theme_classic() + 
  scale_y_continuous(trans = "log10") +  # log-transform y-axis
   labs(x = "sample type",
       y = "DNA concentration (ng/uL)",
       fill = "low reads?") 
dna_conc_sample_type
```

```{r}
ggsave("figures/dna_conc_sample_type.png", plot = dna_conc_sample_type, width = 6, height = 4, dpi = 300)
```



```{r}
dna_conc_dissim <- rep_metadata %>%
  filter(low_reads == "no") %>%
  ggplot(aes(x = tank_replicate_outlier, y = DNA_conc)) +
  geom_boxplot() + 
  theme_classic() + 
  scale_y_continuous(trans = "log10") +  # log-transform y-axis
   labs(x = "tank dissimilarity?",
       y = "DNA concentration (ng/uL)") 
dna_conc_dissim
```

```{r}
ggsave("figures/dna_conc_dissim.png", plot = dna_conc_dissim, width = 6, height = 4, dpi = 300)
```


```{r}
rep_metadata %>%
  #filter(low_reads == "no") %>%
  group_by(sample_type, low_reads) %>%
  summarize(count = n())

blanks <- rep_metadata %>%
  filter(low_reads == "no") %>%
  filter(sample_type == "field_blank")
```



- simple, raw, and mock results all here
```{r}
"joined_community_estimates.csv"
```


- removed duplicate samples and combined with dissimilarity measures
```{r}
"tank_mock_estimates.csv"
"tank_simple_estimates.csv"
"tank_raw_estimates.csv"
```



