---
title: "a look at the quantitative metabarcoding results - aquaria samples"
author: "Kimberly Ledger"
date: "2023-10-06"
output: html_document
---

load libraries
```{r}
library(tidyverse)
library(data.table)
library(gridExtra)
library(ggsci)
select <- dplyr::select
```

# Read in the stan model outputs 
```{r}
#load("/home/kimberly.ledger/gadid_aquaria/gadid_mb_aquaria/quant_mb_outputs/raw_20231101.Rdata")
load("/home/kimberly.ledger/gadid_aquaria/gadid_mb_aquaria/quant_mb_outputs/raw_20240313.Rdata.Rdata")
raw <- Output

#load("/home/kimberly.ledger/gadid_aquaria/gadid_mb_aquaria/quant_mb_outputs/mock_20240313.Rdata")
load("/home/kimberly.ledger/gadid_aquaria/gadid_mb_aquaria/quant_mb_outputs/mock_20231102.Rdata")
#mock1 <- Output 
mock1 <- Output_mock
```

```{r}
#########################################################
# Raw estimates from Reads
#########################################################
# summarize raw estimates from reads for each species.
raw.raw <- raw$env %>% 
  group_by(community,Cycles,tech_rep) %>%
  mutate(sum.ng = sum(start_conc_ng),
         true.prop = start_conc_ng / sum.ng) %>%
  ungroup() %>%
  group_by(Species,community,Cycles,true.prop) %>%
  group_by(Species,community,Cycles) %>%
  summarize(simple.Mean=mean(propReads),
            simple.N = length(tech_rep)) %>%
  replace_na(list(raw.Mean=0,raw.SD=0,raw.SE=0))

# extract predicted proportions from the posterior
COM <- data.frame(community = levels(raw$env$community %>% as.factor()))
COM$comm_idx <- 1:nrow(COM)
SP  <- raw$env %>% distinct(Species,sp_idx) %>% as.data.frame()

# These are the predicted intercepts for the posteriors
beta_posterior <- raw$stanMod_summary[["int_samp_small"]][, c(1,4:8)]
colnames(beta_posterior) <- paste0("raw.",substr(colnames(beta_posterior),1,nchar(colnames(beta_posterior))-1))
colnames(beta_posterior)[1] <- "raw.mean"
beta_posterior <- as.data.frame(beta_posterior)

raw.post <-expand.grid(comm_idx = COM$comm_idx,sp_idx =SP$sp_idx) %>% 
  arrange(comm_idx,sp_idx) %>% 
  left_join(.,COM) %>% 
  left_join(.,SP) %>% 
  bind_cols(.,beta_posterior)

# Combine the raw estimates and posterior estimates
raw.all <- full_join(raw.raw,raw.post)
```

```{r}
#########################################################
# Mock1
#########################################################
# summarize raw estimates from reads for each species.
mock1.raw <- mock1$env %>% group_by(community,Cycles,tech_rep) %>%
      mutate(sum.ng = sum(start_conc_ng),
             true.prop = start_conc_ng / sum.ng) %>%
      ungroup() %>%
      group_by(Species,community,Cycles,true.prop) %>%
  summarize(simple.Mean=mean(propReads),
            simple.N = length(tech_rep)) %>%
  replace_na(list(raw.Mean=0,raw.SD=0,raw.SE=0))

# extract predicted proportions from the posterior
COM <- data.frame(community = levels(mock1$env$community %>% as.factor()))
COM$comm_idx <- 1:nrow(COM)
SP  <- mock1$env %>% distinct(Species,sp_idx) %>% as.data.frame()

# These are the predicted intercepts for the posteriors
beta_posterior <- mock1$stanMod_summary[["int_samp_small"]][, c(1,4:8)]
colnames(beta_posterior) <- paste0("mock1.",substr(colnames(beta_posterior),1,nchar(colnames(beta_posterior))-1))
colnames(beta_posterior)[1] <- "mock1.mean"
beta_posterior <- as.data.frame(beta_posterior)

mock1.post <-expand.grid(comm_idx = COM$comm_idx,sp_idx =SP$sp_idx) %>% 
    arrange(comm_idx,sp_idx) %>% 
    left_join(.,COM) %>% 
    left_join(.,SP) %>% 
    bind_cols(.,beta_posterior)

# Combine the raw estimates and posterior estimates
mock1.all <- full_join(mock1.raw,mock1.post)
```

```{r}
# Combine mock results with raw reads.
result.dat <- left_join(mock1.all,raw.all) %>%
  filter(Species != "Arctogadus glacialis") %>%   #remove species that are not in tanks 
  filter(Species != "Eleginus gracilis") %>%
  filter(Species != "Microgadus proximus") 
```

get the tank IDs 
```{r}
metadata <- read.csv("/home/kimberly.ledger/gadid_aquaria/gadid_mb_aquaria/gadid_aquariaDBO_metadata.csv") %>%
  filter(project == "aquaria") %>%
  select(!station_ID) %>%
  select(!location1:time_of_day) %>%
  select(!Sample_ID) %>%
  select(!pcr_replicate) %>%
  select(!project) %>%
  dplyr::rename(community = extraction_ID) %>%
  unique()
```


join read proportions with metadata
```{r}
result.join <- result.dat %>%
  left_join(metadata)
```

just looking at one tank first
```{r}
tank_40 <- result.join %>%
  filter(tank_ID == 40)
```

plot community 1 - posterior estimates
```{r}
post_40 <- ggplot(tank_40) + # %>% 
          geom_errorbar(aes(x=Species,ymin=raw.2.5,ymax=raw.97.5,color=Species),width=0) +
          geom_errorbar(aes(x=Species,ymin=raw.25,ymax=raw.75,color=Species),width=0, linewidth=2) +        
          geom_point(aes(x=Species,y=raw.mean,color=Species),shape =21,fill="black",size=2) +
          #scale_y_continuous(NULL,expand=c(0,NA),limits=c(0,NA),breaks=yBREAKS) +
          #scale_shape_manual(values=c(21,22,24)) +
          #scale_fill_manual(values=col.val,"Community") +
          #scale_color_manual(values=col.val,"Calibration") +
          #scale_x_continuous(NULL,breaks=xBREAKS$comm.idx,labels=xBREAKS$community,limits=c(NA,3.3)) +
          ylim(0, 1) +
          facet_wrap(~community) + 
          ylab("proportion of reads") +
          theme_classic() +
          theme(#legend.position = c(0.4755,0.85),
                legend.position = "right",
                plot.margin = margin(0,0,0,0.85,"lines"),
                legend.key.size=unit(0.1,'lines'),
                legend.text=element_text(size=7),
                legend.title=element_text(size=9))
          
post_40
```

plot community 1 - posterior estimates
```{r}
simple_40 <- ggplot(tank_40) + # %>% 
          geom_point(aes(x=Species,y=simple.Mean,color=Species),shape =21,fill="black",size=2) +
          facet_wrap(~community) + 
          ylab("proportion of reads") +
          ylim(0, 1) +
          theme_classic() +
          theme(#legend.position = c(0.4755,0.85),
                legend.position = "right",
                plot.margin = margin(0,0,0,0.85,"lines"),
                legend.key.size=unit(0.1,'lines'),
                axis.text.x = element_blank(),
                legend.text=element_text(size=7),
                legend.title=element_text(size=9))
          
simple_40
```


plot mean raw and mean mock-adjusted for a tank 
```{r}
temp1 <- result.join %>%
  filter(tank_ID == 40) %>%
  dplyr::select(Species, community, simple.Mean, mock1.mean, raw.mean)

temp2 <- temp1[,c(2:6)] %>%
  pivot_longer(cols = c(3:5), names_to = "method", values_to = "prop")

# get rid of reference species denotion
temp2 <- temp2 %>%
  mutate(SP= ifelse(grepl("zRefSpecies_",Species),
  substr(Species,13,nchar(Species)),
  as.character(Species)))

plot_prop <- temp2 %>%
  ggplot(aes(x=community, y=prop, fill=SP)) +
  geom_bar(stat = "identity") + 
  theme_bw() +
  facet_grid(~method, scales = 'free') +
  labs(
    y = "proportion of sequencing reads",
    x = "method",
    title = "tank 40") + 
  theme(
    axis.text.x = element_text(angle = 90, hjust = 0.95),
    legend.text = element_text(size = 8),
    legend.key.size = unit(0.3, "cm"),
    legend.position = "right",
    legend.title = element_blank()
  )

plot_prop
```

plot mean raw and mean mock-adjusted for a tank 
```{r}
temp1 <- result.join %>%
  filter(tank_ID == 17) %>%
  dplyr::select(Species, community, simple.Mean, mock1.mean, raw.mean)

temp2 <- temp1[,c(2:6)] %>%
  pivot_longer(cols = c(3:5), names_to = "method", values_to = "prop")

# get rid of reference species denotion
temp2 <- temp2 %>%
  mutate(SP= ifelse(grepl("zRefSpecies_",Species),
  substr(Species,13,nchar(Species)),
  as.character(Species)))

plot_prop <- temp2 %>%
  ggplot(aes(x=community, y=prop, fill=SP)) +
  geom_bar(stat = "identity") + 
  theme_bw() +
  facet_grid(~method, scales = 'free') +
  labs(
    y = "proportion of sequencing reads",
    x = "method",
    title = "tank 17") + 
  theme(
    axis.text.x = element_text(angle = 90, hjust = 0.95),
    legend.text = element_text(size = 8),
    legend.key.size = unit(0.3, "cm"),
    legend.position = "right",
    legend.title = element_blank()
  )

plot_prop
```

plot mean raw and mean mock-adjusted for a tank 
```{r}
temp1 <- result.join %>%
  filter(tank_ID == 71) %>%
  dplyr::select(Species, community, simple.Mean, mock1.mean, raw.mean)

temp2 <- temp1[,c(2:6)] %>%
  pivot_longer(cols = c(3:5), names_to = "method", values_to = "prop")

# get rid of reference species denotion
temp2 <- temp2 %>%
  mutate(SP= ifelse(grepl("zRefSpecies_",Species),
  substr(Species,13,nchar(Species)),
  as.character(Species)))

plot_prop <- temp2 %>%
  ggplot(aes(x=community, y=prop, fill=SP)) +
  geom_bar(stat = "identity") + 
  theme_bw() +
  facet_grid(~method, scales = 'free') +
  labs(
    y = "proportion of sequencing reads",
    x = "method",
    title = "tank 71") + 
  theme(
    axis.text.x = element_text(angle = 90, hjust = 0.95),
    legend.text = element_text(size = 8),
    legend.key.size = unit(0.3, "cm"),
    legend.position = "right",
    legend.title = element_blank()
  )

plot_prop
```

plot mean raw and mean mock-adjusted for a tank 
```{r}
temp1 <- result.join %>%
  filter(tank_ID == 38) %>%
  dplyr::select(Species, community, simple.Mean, mock1.mean, raw.mean)

temp2 <- temp1[,c(2:6)] %>%
  pivot_longer(cols = c(3:5), names_to = "method", values_to = "prop")

# get rid of reference species denotion
temp2 <- temp2 %>%
  mutate(SP= ifelse(grepl("zRefSpecies_",Species),
  substr(Species,13,nchar(Species)),
  as.character(Species)))

plot_prop <- temp2 %>%
  ggplot(aes(x=community, y=prop, fill=SP)) +
  geom_bar(stat = "identity") + 
  theme_bw() +
  facet_grid(~method, scales = 'free') +
  labs(
    y = "proportion of sequencing reads",
    x = "method",
    title = "tank 38") + 
  theme(
    axis.text.x = element_text(angle = 90, hjust = 0.95),
    legend.text = element_text(size = 8),
    legend.key.size = unit(0.3, "cm"),
    legend.position = "right",
    legend.title = element_blank()
  )

plot_prop
```

okay, now let me adapt the code from the decontamination pipeline to see what tank replicates, if any, have high dissimilarity 

instead of filtering by the biological replicates, i need to filter by the unique extractions.  the 'dist_to_centroid' function later on doesn't like the different number of replicates or that there are so many for a few.. not sure exactly. 

how many replicates does each tank have? 
```{r}
result.join %>%
  group_by(tank_ID) %>%
  summarise(nrep = n_distinct(community)) %>%
  #filter(nrep == 2)  # there are two (51 and 59) 
  filter(nrep == 1) # there is one 
```

remove the tanks with only one pcr rep from the data frame
```{r}
result.temp <- result.join %>%
  filter(tank_ID != 54)
```

first, i'll calculate an eDNA index for the mock1.mean (this normalization doesn't really change the prop that much...)
```{r}
normalized <- result.temp %>%
  dplyr::group_by(Species) %>%
  mutate(Colmax = max(mock1.mean, na.rm = TRUE),     ## update with simple, mock, or raw mean
         Normalized_reads = mock1.mean/Colmax)

#add a new sample id column that also includes the tank - will use this for dissimilarity measures
normalized <- normalized %>%
  unite(tank_community, tank_ID, community, sep = "_", remove = FALSE)
```


```{r}
library(vegan)

tibble_to_matrix <- function (tb) {
  
  tb %>%
  #normalized %>%
    group_by(Species, tank_community) %>% 
    summarise(nReads = sum(Normalized_reads)) %>% 
    spread ( key = "Species", value = "nReads", fill = 0) %>%
    ungroup() -> matrix_1
    samples <- pull (matrix_1, tank_community)
    matrix_1[,-1] -> matrix_1
    data.matrix(matrix_1) -> matrix_1
    dimnames(matrix_1)[[1]] <- samples
    vegdist(matrix_1) -> matrix_1
}
```


```{r}
all.distances.full <- tibble_to_matrix(normalized)

# Do all samples have a name?
summary(is.na(names(all.distances.full)))
```

make the pairwise distances a long table
```{r}
as_tibble(subset(melt(as.matrix(all.distances.full)))) -> all.distances.melted

# Any major screw ups
summary(is.na(all.distances.melted$value))

# Now, create a three variables for all distances, they could be PCR replicates, BIOL replicates, or from the same site

all.distances.melted %>%
  separate (Var1, into = c("Tank1", "community1"), sep = "\\_", remove = FALSE) %>%
  separate (Var2, into = c("Tank2", "community2"), sep = "\\_", remove = FALSE) %>%
  mutate (Distance.type = case_when(community1 == community2 ~ "Biological Replicates",
                                      Tank1 == Tank2 ~ "Same Tank",
                                      TRUE ~ "Different Tank"
                                     )) %>%
  dplyr::select(Sample1 = Var1, Sample2 = Var2 , value , Distance.type) %>%
  filter (Sample1 != Sample2) -> all.distances.to.plot

# Checking all went well

sapply(all.distances.to.plot, function(x) summary(is.na(x)))
```

```{r}
all.distances.to.plot$Distance.type <- all.distances.to.plot$Distance.type  %>% fct_relevel("Biological Replicates", "Same Tank")

myplot <- ggplot (all.distances.to.plot) +
  geom_histogram (aes (fill = Distance.type, x = value, after_stat(ndensity)), position = "dodge",  alpha = 0.9, bins = 50) +
  facet_wrap( ~ Distance.type) +
  theme_bw() +
  labs (x = "Pairwise dissimilarity", y = "Density" ,
        Distance.type = "Distance") +
    guides (fill = "none")
myplot
```

cool, samples from the same tank are usually quite similar. 

```{r}
#ggsave("figures/tank_pairwise_dissimiarlity.png", plot = myplot, width = 8.5, height = 5, dpi = 300)
```


summarize pairwise dissimilarity results 
```{r}
all.distances.to.plot %>%
  group_by(Distance.type) %>%
  summarize(Mean = mean(value),
            SD=sd(value),
                q.025 = quantile(value,probs=0.025),
                #q.05 = quantile(value,probs=0.05),
                #q.25 = quantile(value,probs=0.25),
                #q.75 = quantile(value,probs=0.75),
                #q.95 = quantile(value,probs=0.95),
                q.975 = quantile(value,probs=0.975))
```

which samples/tanks are about the 0.25 dissimilarity 
```{r}
pairwise <- all.distances.to.plot %>%
  filter(Distance.type == "Same Tank") %>%
  filter(value > 0.25)
pairwise
```

```{r}
write.csv(pairwise, "pairwise_mock_tank_dissimilarity.csv", row.names = F)
```


next i will follow what was done here:  (https://github.com/ramongallego/eDNA.and.Ocean.Acidification.Gallego.et.al.2020/blob/master/Scripts/Denoising.all.runs.Rmd) and instead of choosing outliers based on the pairwise distances, we can do a similar thing using the distance to centroid. 

now identify and discard outliers 
```{r message=FALSE, warning=FALSE}
normalized %>%
  group_by(tank_ID) %>% nest() -> nested.cleaning 

nested.cleaning %>% 
  mutate(matrix = map(data, tibble_to_matrix)) -> nested.cleaning

nested.cleaning %>% mutate(ncomparisons = map(matrix, length)) -> nested.cleaning
```

```{r}
dist_to_centroid <- function (x,y) {
  
  #biol <- rep(y, dim(x)[[1]])
  biol <- rep(y, length(x))
  
  if (length(biol) == 1) {
    output = rep(x[1]/2,2)
    names(output) <- attr(x, "Labels")
  }else{ 
    
  dispersion <- betadisper(x, group = biol)
  output = dispersion$distances
  }
  output
    }
```

```{r}
nested.cleaning.temp <- nested.cleaning %>% 
  mutate(distances = map2(matrix, tank_ID, dist_to_centroid))

all_distances <- nested.cleaning.temp %>%
  unnest_longer(distances) %>%
  dplyr::select(tank_ID, distances_id, distances)

hist(all_distances$distances)
```

calculate normal distribution of distances to centroid
```{r}
normparams <- MASS::fitdistr(all_distances$distances, "normal")$estimate                                      
probs <- pnorm(all_distances$distances, normparams[1], normparams[2])
outliers_centroid <- which(probs>0.95)

discard_centroid <- all_distances$distances_id[outliers_centroid]
discard_centroid
```

okay there are a few more tanks included here than than before. i'll go back up to the code earlier and plot some of them.   
11_e02158, 41_e02158, is a bit different than the other reps, but not by much 

```{r}
#write.csv(discard_centroid, "centroid_mock_tank_dissimilarity.csv", row.names = F)
```

```{r}
#write.csv(result.join, "joined_community_estimates.csv", row.names = F)
```

