---
title: "summarize metabarcoding tank proportions"
author: "Kimberly Ledger"
date: "2023-11-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

load libraries
```{r}
library(tidyverse)
library(dplyr)
```

read in sample sheets that includes all libraries (sample replicates) included in sequencing to do complete tracking of replicates through the analysis workflow
```{r}
metadata <- read.csv("/home/kimberly.ledger/gadid_aquaria/gadid_mb_aquaria/gadid_aquariaDBO_metadata.csv") %>%
  filter(project == "aquaria") %>%
  dplyr::select(!c(location1:time_of_day)) %>%
  dplyr::select(!station_ID)

#illumina output changed "_" to "-"
metadata$Sample_ID <- gsub("_", "-", metadata$Sample_ID) 
```

remove the samples from miseq run B that were also included in miseq run A (as i did during the data processing)
```{r}
both_runs_extID <- c("e02362", "e02363", "e02366", "e02370", "e02372", "e02373", "e02374", "e02375", "e02380")

both_runs <- metadata %>%
  filter(extraction_ID %in% both_runs_extID)

both_runs_A <- both_runs %>%
  filter(MiSeq_run == "A")

metadata <- metadata %>%
  filter(!extraction_ID %in% both_runs_extID) %>%
  bind_rows(both_runs_A) 
```

reformat metadata that removes the sample_ID and adds a column for the number of pcr replicates for each extraction_ID
```{r}
metadata_2 <- metadata %>%
  group_by(extraction_ID) %>%
  mutate(N.pcr.reps = n()) %>%
  dplyr::select(!Sample_ID) %>%
  dplyr::select(!pcr_replicate) %>%
  dplyr::select(!project) %>%
  unique()
```



read in the extraction/biological replicate/community level estimates of read proportions (simple mean, raw mean and mock-community adjusted reads)

- okay, before i ran the quant metabarcoding model, the samples that were in both miseq runs got duplicated because i did not remove the B run from metadata... only the reads from the A run were processed, but the export say both A and B... need to do some extra thing to fix this. 
```{r}
quant_results_to_replace <- read.csv("joined_community_estimates.csv") %>%
  filter(community %in% both_runs_extID) %>%
  filter(!MiSeq_run == "B")
  
quant_results <- read.csv("joined_community_estimates.csv") %>%
  filter(!community %in% both_runs_extID) %>%
  bind_rows(quant_results_to_replace) %>%
  dplyr::select(Species, community, simple.N, simple.Mean, raw.mean, mock1.mean) %>%
  mutate(Species = ifelse(Species == "zRefSpecies_Gadus macrocephalus", "Gadus macrocephalus", Species)) %>%
  mutate(simple.Mean = ifelse(is.na(simple.Mean), 0, simple.Mean))

quant_wide <- quant_results %>%
  pivot_wider(
    names_from = "Species",
    values_from = c("simple.Mean", "raw.mean", "mock1.mean", values_fill = 0)
  ) %>%
  rename(extraction_ID = community)
```


join the two dataframes
```{r}
join <- metadata_2 %>%
  left_join(quant_wide, by = "extraction_ID")
```

okay, now can i indicate which extraction IDs had outlier pcr-replicate dissimilarity (using taxa level test)
```{r}
pcr_dissimilarity <- read.csv("pcr_dissimilarity_outliers.csv") %>%
  dplyr::select(!X)

#these are indicated by tank_altID-replicate - to seperate
pcr_dissimilarity <- pcr_dissimilarity %>%
  separate(col = discard_centroid, into = c("tank_ID", "alt_ID"))
```

```{r}
join <- join %>%
  mutate(pcr_dissim = ifelse(alt_ID %in% pcr_dissimilarity$alt_ID, "yes", "no")) %>%
  mutate(pcr_dissim = ifelse(is.na(simple.N), NA, pcr_dissim))
```


now add in the tank dissimilarity estimates
```{r}
mock_dis <- read.csv("centroid_mock_tank_dissimilarity.csv")

#okay these have all the same IDs, with the exception of one missing from the simple... will just use the raw_dis object then

mock_dis <- mock_dis %>%
  separate(col = x, into = c("tank_ID", "extraction_ID"))
```

```{r}
join <- join %>%
  mutate(tank_dissim = ifelse(extraction_ID %in% mock_dis$extraction_ID, "yes", "no")) %>%
  mutate(tank_dissim = ifelse(is.na(simple.N), NA, tank_dissim)) 
```

export a similified verion of this to share with collaborators 
```{r}
join_simple <- join %>%
  dplyr::select(!c(extraction_plate:collection_day)) %>%
  arrange(tank_ID)
```

```{r}
#write.csv(join_simple, "sample_estimates.csv", row.names = F)
```


```{r}
join_simple %>%
  filter(is.na(simple.N)) %>%
  group_by(tank_ID) %>%
  summarize(count = n()) %>%
  filter(count == 3)
```


now summarize the proportions at the tank level - will just do this with the mock1 estimates for now... 
```{r}
#to keep the T tanks separate, i need to modify their tank names
join_simple[174,3] <- "T_118"
join_simple[188,3] <- "T_55"
join_simple[191,3] <- "T_188"

tanks <- join_simple %>%
  select(!`simple.Mean_Boreogadus saida`:`raw.mean_Gadus macrocephalus`) %>%
  pivot_longer(cols = `mock1.mean_Boreogadus saida`:`mock1.mean_Gadus macrocephalus`, names_to = "Species", values_to = "mock1.mean") %>%
  mutate(Species = ifelse(Species == "mock1.mean_Boreogadus saida", "Boreogadus saida", Species)) %>%
  mutate(Species = ifelse(Species == "mock1.mean_Gadus chalcogrammus", "Gadus chalcogrammus", Species)) %>%
  mutate(Species = ifelse(Species == "mock1.mean_Gadus macrocephalus", "Gadus macrocephalus", Species)) %>%
  filter(!is.na(simple.N)) %>%
  group_by(tank_ID, Species) %>%
  summarise(n.simple.N = sum(simple.N),
            n.sample.dissim = sum(pcr_dissim == "yes", na.rm = TRUE),
            n.tank.dissim = sum(tank_dissim == "yes", na.rm = TRUE),
            Mean = mean(mock1.mean),
            SD=sd(mock1.mean),
                q.025 = quantile(mock1.mean,probs=0.025),
                #q.05 = quantile(mock1.mean,probs=0.05),
                #q.25 = quantile(mock1.mean,probs=0.25),
                #q.75 = quantile(mock1.mean,probs=0.75),
                #q.95 = quantile(mock1.mean,probs=0.95),
                q.975 = quantile(mock1.mean,probs=0.975))  
```

```{r}
write.csv(tanks, "aquaria_DBO/tank_mock_estimates.csv", row.names = F)
```



```{r}
tank_removeoutliers <- join_simple %>%
  select(!`simple.Mean_Boreogadus saida`:`raw.mean_Gadus macrocephalus`) %>%
  pivot_longer(cols = `mock1.mean_Boreogadus saida`:`mock1.mean_Gadus macrocephalus`, names_to = "Species", values_to = "mock1.mean") %>%
  mutate(Species = ifelse(Species == "mock1.mean_Boreogadus saida", "Boreogadus saida", Species)) %>%
  mutate(Species = ifelse(Species == "mock1.mean_Gadus chalcogrammus", "Gadus chalcogrammus", Species)) %>%
  mutate(Species = ifelse(Species == "mock1.mean_Gadus macrocephalus", "Gadus macrocephalus", Species)) %>%
  filter(!is.na(simple.N)) %>%
  filter(tank_dissim == "no") %>%
  group_by(tank_ID, Species) %>%
  summarise(n.simple.N = sum(simple.N),
            n.sample.dissim = sum(pcr_dissim == "yes", na.rm = TRUE),
            #n.tank.dissim = sum(tank_dissim == "yes", na.rm = TRUE),
            Mean = mean(mock1.mean),
            SD=sd(mock1.mean),
                q.025 = quantile(mock1.mean,probs=0.025),
                #q.05 = quantile(mock1.mean,probs=0.05),
                #q.25 = quantile(mock1.mean,probs=0.25),
                #q.75 = quantile(mock1.mean,probs=0.75),
                #q.95 = quantile(mock1.mean,probs=0.95),
                q.975 = quantile(mock1.mean,probs=0.975))  
```

```{r}
#write.csv(tank_removeoutliers, "aquaria_DBO/tank_mock_estimates_no_outliers.csv", row.names = F)
```


plot 
```{r}
tank_prop %>%
  filter(tank_ID == 10)
```

```{r}
tank_list <- c(11:30)

tank_prop %>%
  filter(tank_ID %in% tank_list) %>%
  ggplot() +
      geom_errorbar(aes(x=tank_ID,ymin=q.025,ymax=q.975,color=Species),width=0) +
      geom_errorbar(aes(x=tank_ID,ymin=q.25,ymax=q.75,color=Species),size=2,width=0) +        
      geom_point(aes(x=tank_ID,y=Mean,color=Species),size=3) +
      ylab("mean proportion") +
      theme_bw()
```
