---
title: "exploring mb estimates v true"
author: "Kimberly Ledger"
date: "2023-11-22"
output: html_document
---

this code generates lots of figures to explore the data.  see 8_mb_aquarium_plots.Rmd for final figure plots.  

```{r}
library(tidyverse)
library(ggplot2)
select <- dplyr::select
```

read in tank estimates
```{r}
tanks <- read.csv("/home/kimberly.ledger/gadid_aquaria/gadid_mb_aquaria/tank_mock_estimates.csv") %>%
  mutate(Species = ifelse(Species == "Boreogadus saida", "AC", Species),
         Species = ifelse(Species == "Gadus chalcogrammus", "WP", Species),
         Species = ifelse(Species == "Gadus macrocephalus", "PC", Species)) %>%
  pivot_wider(names_from = Species, values_from = c(6:9))
```

read in tank metadata 
```{r}
metadata <-  read.csv("/home/kimberly.ledger/gadid_aquaria/gadid_mb_aquaria/TankData_MBmodified.csv") %>% 
  filter(groupID != "T") %>%  #remove tap water controls
  select(!sampleID) %>% #remove sample id because estimates are at the tank-level
  unique() %>%
  rename(tank_ID = groupID)
```

join - this removes the few T tanks with eDNA reads. okay for now but will need to report those somewhere... 
```{r}
all <- metadata %>%
  left_join(tanks, by = "tank_ID")
```

```{r}
all$n.tank.dissim <- as.factor(all$n.tank.dissim)
all$n.sample.dissim <- as.factor(all$n.sample.dissim)
```

tanks with no fish
```{r}
all %>% 
  filter(Fish_pres == 0)
```


# PLOTS!!!!!

### TANK DISSIMILARITY 

```{r}
all %>%
  ggplot() + 
  geom_errorbar(aes(x=AC_Fbiomass,ymin=q.025_AC,ymax=q.975_AC,color=n.tank.dissim),width=0) +
  geom_point(aes(x=AC_Fbiomass,y=Mean_AC,color=n.tank.dissim)) +
  scale_color_manual(values = c("0" = "black", "1" = "orange", "2" = "red"))
```

```{r}
all %>%
  ggplot() + 
  geom_errorbar(aes(x=PC_Fbiomass,ymin=q.025_PC,ymax=q.975_PC,color=n.tank.dissim),width=0) +
  geom_point(aes(x=PC_Fbiomass,y=Mean_PC,color=n.tank.dissim)) +
  scale_color_manual(values = c("0" = "black", "1" = "orange", "2" = "red"))
```

```{r}
all %>%
  ggplot() + 
  geom_errorbar(aes(x=WP_Fbiomass,ymin=q.025_WP,ymax=q.975_WP,color=n.tank.dissim),width=0) +
  geom_point(aes(x=WP_Fbiomass,y=Mean_WP,color=n.tank.dissim)) +
  scale_color_manual(values = c("0" = "black", "1" = "orange", "2" = "red"))
```

okay, excluding tanks where there is 2 dissimilar estimates seems like a good idea 

### SAMPLE DISSIMILARITY 

```{r}
all %>%
  filter(n.tank.dissim != "2") %>%
  ggplot() + 
  geom_errorbar(aes(x=AC_Fbiomass,ymin=q.025_AC,ymax=q.975_AC,color=n.sample.dissim),width=0) +
  geom_point(aes(x=AC_Fbiomass,y=Mean_AC,color=n.sample.dissim)) +
  scale_color_manual(values = c("0" = "black", "1" = "orange", "2" = "red", "3" = "purple"))
```

```{r}
all %>%
  filter(n.tank.dissim != "2") %>%
  ggplot() + 
  geom_errorbar(aes(x=PC_Fbiomass,ymin=q.025_PC,ymax=q.975_PC,color=n.sample.dissim),width=0) +
  geom_point(aes(x=PC_Fbiomass,y=Mean_PC,color=n.sample.dissim)) +
  scale_color_manual(values = c("0" = "black", "1" = "orange", "2" = "red", "3" = "purple"))
```

```{r}
all %>%
  filter(n.tank.dissim != "2") %>%
  ggplot() + 
  geom_errorbar(aes(x=WP_Fbiomass,ymin=q.025_WP,ymax=q.975_WP,color=n.sample.dissim),width=0) +
  geom_point(aes(x=WP_Fbiomass,y=Mean_WP,color=n.sample.dissim)) +
  scale_color_manual(values = c("0" = "black", "1" = "orange", "2" = "red", "3" = "purple"))
```

sample dissimilarity on it's own doesn't seem like something i need to worry about/remove. 

### DAYS SINCE FISH IN TANK 

```{r}
all %>%
  filter(n.tank.dissim != "2") %>%
  ggplot() + 
  geom_errorbar(aes(x=AC_Fbiomass,ymin=q.025_AC,ymax=q.975_AC,color=DaysSinceAC),width=0) +
  geom_point(aes(x=AC_Fbiomass,y=Mean_AC,color=DaysSinceAC)) + 
  scale_color_continuous(low = "blue", high = "red")

all %>%
  filter(n.tank.dissim != "2") %>%
  filter(DaysSinceAC > 0) %>%
  ggplot() + 
  geom_errorbar(aes(x=tank_ID,ymin=q.025_AC,ymax=q.975_AC,color=DaysSinceAC),width=0) +
  geom_point(aes(x=tank_ID,y=Mean_AC,color=DaysSinceAC)) + 
  scale_color_continuous(low = "blue", high = "red")
```

```{r}
all %>%
  filter(n.tank.dissim != "2") %>%
  ggplot() + 
  geom_errorbar(aes(x=PC_Fbiomass,ymin=q.025_PC,ymax=q.975_PC,color=DaysSincePC),width=0) +
  geom_point(aes(x=PC_Fbiomass,y=Mean_PC,color=DaysSincePC)) + 
  scale_color_continuous(low = "blue", high = "red")

all %>%
  filter(n.tank.dissim != "2") %>%
  filter(DaysSincePC > 0) %>%
  ggplot() + 
  geom_errorbar(aes(x=tank_ID,ymin=q.025_PC,ymax=q.975_PC,color=DaysSincePC),width=0) +
  geom_point(aes(x=tank_ID,y=Mean_PC,color=DaysSincePC)) + 
  scale_color_continuous(low = "blue", high = "red")
```

```{r}
all %>%
  filter(n.tank.dissim != "2") %>%
  ggplot() + 
  geom_errorbar(aes(x=WP_Fbiomass,ymin=q.025_WP,ymax=q.975_WP,color=DaysSinceWP),width=0) +
  geom_point(aes(x=WP_Fbiomass,y=Mean_WP,color=DaysSinceWP)) + 
  scale_color_continuous(low = "blue", high = "red")

all %>%
  filter(n.tank.dissim != "2") %>%
  filter(DaysSinceWP > 0) %>%
  ggplot() + 
  geom_errorbar(aes(x=tank_ID,ymin=q.025_WP,ymax=q.975_WP,color=DaysSinceWP),width=0) +
  geom_point(aes(x=tank_ID,y=Mean_WP,color=DaysSinceWP)) + 
  scale_color_continuous(low = "blue", high = "red")
```

the time since fish were last in a tank doesn't seem to be a big issue, except for WP (residual DNA (probably not after 50 days...) or false positive)

### SUMMARIZE FALSE POSITIVE DETETECTIONS
```{r}
# arctic cod
all %>%
  filter(n.tank.dissim != "2") %>% 
  filter(Mean_AC > 0.005) %>%
  filter(AC_Fbiomass == 0)
```

```{r}
# pacific cod
all %>%
  filter(n.tank.dissim != "2") %>% 
  filter(Mean_PC > 0.005) %>%
  filter(PC_Fbiomass == 0)
```

```{r}
# walleye pollock
all %>%
  filter(n.tank.dissim != "2") %>% 
  filter(Mean_WP > 0.005) %>%
  filter(WP_Fbiomass == 0)
```

### TIME SINCE INTRODUCTION OF FISH INTO TANK 

reformat tank time into days
```{r}
all <- all %>%
  mutate(TankTime_Days = ifelse(TankTime == ">four years", 1460, NA),
         TankTime_Days = ifelse(TankTime == ">three years", 1095, TankTime_Days),
         TankTime_Days = ifelse(TankTime == "1 day", 1, TankTime_Days),
         TankTime_Days = ifelse(TankTime == "2 weeks", 14, TankTime_Days),
         TankTime_Days = ifelse(TankTime == "3 days", 3, TankTime_Days),
         TankTime_Days = ifelse(TankTime == "3 weeks", 21, TankTime_Days),
         TankTime_Days = ifelse(TankTime == "4 days", 4, TankTime_Days),
         TankTime_Days = ifelse(TankTime == "5 days", 5, TankTime_Days),
         TankTime_Days = ifelse(TankTime == "7 days", 7, TankTime_Days), 
         TankTime_Days = ifelse(TankTime == "present in tank for years; over a week in this temp range", 1460, TankTime_Days) )
```


```{r}
all %>%
  filter(n.tank.dissim != "2") %>%
  ggplot() + 
  geom_errorbar(aes(x=AC_Fbiomass,ymin=q.025_AC,ymax=q.975_AC,color=TankTime_Days),width=0) +
  geom_point(aes(x=AC_Fbiomass,y=Mean_AC,color=TankTime_Days)) + 
  scale_color_continuous(low = "blue", high = "red", na.value = "gray", trans = "log")
```

```{r}
all %>%
  filter(n.tank.dissim != "2") %>%
  ggplot() + 
  geom_errorbar(aes(x=PC_Fbiomass,ymin=q.025_PC,ymax=q.975_PC,color=TankTime_Days),width=0) +
  geom_point(aes(x=PC_Fbiomass,y=Mean_PC,color=TankTime_Days)) + 
  scale_color_continuous(low = "blue", high = "red", na.value = "gray", trans = "log")
```

the tank that had 100% PC reads was only empty of PC for one day prior to sampling and had no new fish introduced... 
maybe this tank and others that seem like they were "tests" that tom tossed in but don't really fit our main question can be removed from some analyses/figures later on... 


```{r}
all %>%
  filter(n.tank.dissim != "2") %>%
  ggplot() + 
  geom_errorbar(aes(x=WP_Fbiomass,ymin=q.025_WP,ymax=q.975_WP,color=TankTime_Days),width=0) +
  geom_point(aes(x=WP_Fbiomass,y=Mean_WP,color=TankTime_Days)) + 
  scale_color_continuous(low = "blue", high = "red", na.value = "gray", trans = "log")
```

beside tank 3, that had 100% AC for years but eDNA said it had WP and AC, time of fish in tank (2 days to years) doesn't seem to have an influence on results 


### % abundance (number of fish) 

```{r}
all <- all %>%
  mutate(AC_N = ifelse(is.na(AC_N), 0, AC_N), 
         PC_N = ifelse(is.na(PC_N), 0, PC_N),
         WP_N = ifelse(is.na(WP_N), 0, WP_N),
         Tot_N = AC_N + PC_N + WP_N,
         AC_FN = AC_N/Tot_N,
         PC_FN = PC_N/Tot_N,
         WP_FN = WP_N/Tot_N)
```

```{r}
all %>%
  filter(n.tank.dissim != "2") %>%
  ggplot() + 
  geom_errorbar(aes(x=AC_FN,ymin=q.025_AC,ymax=q.975_AC,color=Tot_N),width=0) +
  geom_point(aes(x=AC_FN,y=Mean_AC,color=Tot_N)) + 
  scale_color_continuous(low = "blue", high = "red", na.value = "gray", trans = "log")
```

```{r}
all %>%
  filter(n.tank.dissim != "2") %>%
  ggplot() + 
  geom_errorbar(aes(x=PC_FN,ymin=q.025_PC,ymax=q.975_PC,color=Tot_N),width=0) +
  geom_point(aes(x=PC_FN,y=Mean_PC,color=Tot_N)) + 
  scale_color_continuous(low = "blue", high = "red", na.value = "gray", trans = "log")
```


```{r}
all %>%
  filter(n.tank.dissim != "2") %>%
  ggplot() + 
  geom_errorbar(aes(x=WP_FN,ymin=q.025_WP,ymax=q.975_WP,color=Tot_N),width=0) +
  geom_point(aes(x=WP_FN,y=Mean_WP,color=Tot_N)) + 
  scale_color_continuous(low = "blue", high = "red", na.value = "gray", trans = "log")
```

proportional abundance appears to be far less correlated with eDNA read proportion compared to proportional biomass  

### AGE 

* all AC were age 4
* all WP were age 7 
* only some variation in PC age 

```{r}
library(ggsci)

pc_biomass <- all %>%
  filter(n.tank.dissim != "2") %>%
  filter(Fish_pres == "1") %>%
  ggplot() + 
  geom_errorbar(aes(x=PC_Fbiomass,ymin=q.025_PC,ymax=q.975_PC,color=PC_age),width=0) +
  geom_point(aes(x=PC_Fbiomass,y=Mean_PC,color=PC_age)) + 
  scale_color_material("green") + 
  #geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black")  + 
  geom_smooth(aes(x=PC_Fbiomass,y=Mean_PC), color = "black", linetype = "dashed", method = "lm", se = FALSE) +
  theme_classic() + 
  labs(x = "biomass proportion",
       y = "eDNA read proportion") + 
  theme(legend.position = "none")
pc_biomass
```

```{r}
pc_abund <- all %>%
  filter(n.tank.dissim != "2") %>%
  filter(Fish_pres == "1") %>%
  ggplot() + 
  #geom_errorbar(aes(x=PC_FN,ymin=q.025_PC,ymax=q.975_PC,color=PC_age),width=0) +
  geom_errorbar(aes(x=PC_FN,ymin=q.025_PC,ymax=q.975_PC),width=0, color = "#79af97") +
  #geom_point(aes(x=PC_FN,y=Mean_PC,color=PC_age), size = 3) + 
  geom_point(aes(x=PC_FN,y=Mean_PC), size = 3, color = "#79af97") + 
  scale_color_material("green") + 
  #geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black")  + 
  geom_smooth(aes(x=PC_FN,y=Mean_PC), color = "black", linetype = "dashed", method = "lm", se = FALSE) +
  theme_classic() + 
  labs(x = "proportional abundance",
       y = "eDNA read proportion") + 
  #geom_segment(aes(x = 0.25, y = 0.93333, xend = 0.45, yend = 0.45), linetype = "dotted", color = "red") + # Perpendicular line
  theme(legend.position = "none",
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14))

pc_abund
```


```{r}
#ggsave("figures/PC_propabundance_1color.png", plot = pc_abund, width = 7.5, height = 6, dpi = 300)
```


compute the perpendicular distance of each point to the line
```{r}
dist_data <- all %>%
  filter(n.tank.dissim != "2") %>%
  filter(Fish_pres == "1")

# Calculate distances from points to the 1:1 line
dist_data$distances_abundance <- dist_data$Mean_PC - dist_data$PC_FN
dist_data$distances_biomass <- dist_data$Mean_PC - dist_data$PC_Fbiomass

dist_data <- dist_data %>%
  filter(!is.na(PC_age))
```

```{r}
ggplot(dist_data,aes(x = distances_biomass, fill = PC_age)) + 
  geom_histogram(binwidth = 0.1, aes(group = PC_age)) + 
  theme_bw() + 
  scale_fill_material("green")
```

think of other ways to represent this 

```{r}
dist_data <- dist_data %>%
  mutate(PC_age_group = ifelse(PC_age == 0, "age 0", NA),
         PC_age_group = ifelse(PC_age == 1, "age 1", PC_age_group),
         PC_age_group = ifelse(PC_age == 1.5, "age 1", PC_age_group),
         PC_age_group = ifelse(PC_age == 2, "age 2", PC_age_group),
         PC_age_group = ifelse(PC_age == 7, "age 7", PC_age_group))
```

```{r}
pc_residual_abund <- ggplot(dist_data, aes(y = PC_age_group, x = distances_abundance, fill = PC_age)) + 
  geom_point(shape = 21, cex = 2) + 
  theme_bw() + 
  scale_fill_material("green") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  labs(x = "residual (abundance)",
       y = "Pacific cod age") +
  theme(legend.position = "none")+
  xlim(c(-0.5,0.5))
pc_residual_abund
```

```{r}
pc_residual_biomass <- ggplot(dist_data, aes(y = PC_age_group, x = distances_biomass, fill = PC_age)) + 
  geom_point(shape = 21, cex = 2) + 
  theme_bw() + 
  scale_fill_material("green") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  labs(x = "residual (biomass)",
       y = "Pacific cod age") +
  theme(legend.position = "none") +
  xlim(c(-0.5,0.5))
pc_residual_biomass
```

plot together
```{r, fig.width= 10, fig.height= 6}
library(cowplot)

pc_plot <- plot_grid(pc_biomass, pc_abund, pc_residual_biomass, pc_residual_abund, labels = c("A", "B", "C", "D"), nrow = 2)
pc_plot
```



this is sorta interesting - when looking at proportinal abundance of pcod on x axis, older pcod tend to over represent eDNA proportion and younger pcod tend to under represent eDNA proportions   


### TOTAL BIOMASS

```{r}
all %>%
  filter(n.tank.dissim != "2") %>%
  ggplot() + 
  geom_errorbar(aes(x=AC_Fbiomass,ymin=q.025_AC,ymax=q.975_AC,color=Tot_biomass),width=0) +
  geom_point(aes(x=AC_Fbiomass,y=Mean_AC,color=Tot_biomass)) + 
  scale_color_continuous(low = "blue", high = "red", na.value = "gray", trans = "log")
```


```{r}
all %>%
  filter(n.tank.dissim != "2") %>%
  ggplot() + 
  geom_errorbar(aes(x=PC_Fbiomass,ymin=q.025_PC,ymax=q.975_PC,color=Tot_biomass),width=0) +
  geom_point(aes(x=PC_Fbiomass,y=Mean_PC,color=Tot_biomass)) + 
  scale_color_continuous(low = "blue", high = "red", na.value = "gray", trans = "log")
```


```{r}
all %>%
  filter(n.tank.dissim != "2") %>%
  ggplot() + 
  geom_errorbar(aes(x=WP_Fbiomass,ymin=q.025_WP,ymax=q.975_WP,color=Tot_biomass),width=0) +
  geom_point(aes(x=WP_Fbiomass,y=Mean_WP,color=Tot_biomass)) + 
  scale_color_continuous(low = "blue", high = "red", na.value = "gray", trans = "log")
```


nothing too interesting about total biomass 


### FISH PRESENCE/ABSENCE 

which fish absent tanks had reads? 
```{r}
all %>%
  filter(n.tank.dissim != "2") %>%
  filter(Fish_pres == 0) %>%
  select(DaysSinceAC:DaysSinceWP, Mean_AC:Mean_PC)
```

tank sampled one day after PC removal was all PC reads 

which fish present tanks had no reads? 
```{r}
all %>%
  filter(n.tank.dissim != "2") %>%
  filter(Fish_pres == 1) %>%
  filter(is.na(Mean_AC))
```



### TEMPERATURE 

```{r}
all %>%
  filter(n.tank.dissim != "2") %>%
  ggplot() + 
  geom_errorbar(aes(x=AC_Fbiomass,ymin=q.025_AC,ymax=q.975_AC,color=Temp),width=0) +
  geom_point(aes(x=AC_Fbiomass,y=Mean_AC,color=Temp)) + 
  scale_color_continuous(low = "blue", high = "red", na.value = "gray")
```

```{r}
all %>%
  filter(n.tank.dissim != "2") %>%
  ggplot() + 
  geom_errorbar(aes(x=PC_Fbiomass,ymin=q.025_PC,ymax=q.975_PC,color=Temp),width=0) +
  geom_point(aes(x=PC_Fbiomass,y=Mean_PC,color=Temp)) + 
  scale_color_continuous(low = "blue", high = "red", na.value = "gray")
```


```{r}
all %>%
  filter(n.tank.dissim != "2") %>%
  ggplot() + 
  geom_errorbar(aes(x=WP_Fbiomass,ymin=q.025_WP,ymax=q.975_WP,color=Temp),width=0) +
  geom_point(aes(x=WP_Fbiomass,y=Mean_WP,color=Temp)) + 
  scale_color_continuous(low = "blue", high = "red", na.value = "gray")
```

temperature doesn't seem to have any consistent influence... which maybe is an interesting result?? except most tanks were 4-5C so not really the variability to test anything.  


### TANK - volume, flow rate, and turnover rate vary across tanks 

```{r}
all %>%
  filter(n.tank.dissim != "2") %>%
  ggplot() + 
  geom_errorbar(aes(x=AC_Fbiomass,ymin=q.025_AC,ymax=q.975_AC,color=lab),width=0) +
  geom_point(aes(x=AC_Fbiomass,y=Mean_AC,color=lab))
```

```{r}
all %>%
  filter(n.tank.dissim != "2") %>%
  ggplot() + 
  geom_errorbar(aes(x=PC_Fbiomass,ymin=q.025_PC,ymax=q.975_PC,color=lab),width=0) +
  geom_point(aes(x=PC_Fbiomass,y=Mean_PC,color=lab))
```


```{r}
all %>%
  filter(n.tank.dissim != "2") %>%
  ggplot() + 
  geom_errorbar(aes(x=WP_Fbiomass,ymin=q.025_WP,ymax=q.975_WP,color=lab),width=0) +
  geom_point(aes(x=WP_Fbiomass,y=Mean_WP,color=lab))
```

the lab tank used for each mixture doesn't seem to have any particular influence 


### some models

```{r}
final <- all %>%
  filter(n.tank.dissim != "2") %>%
  filter(Fish_pres == 1)
```


arctic cod 
```{r}
AC <- lm(AC_Fbiomass ~ Mean_AC, data = final)
summary(AC)

AC_N <- lm(AC_FN ~ Mean_AC, data = final)
summary(AC_N)
```

pacific cod 
```{r}
PC <- lm(PC_Fbiomass ~ Mean_PC, data = final)
summary(PC)

PC_N <- lm(PC_FN ~ Mean_PC, data = final)
summary(PC_N)
```

walleye pollock
```{r}
WP <- lm(WP_Fbiomass ~ Mean_WP, data = final)
summary(WP)

WP_N <- lm(WP_FN ~ Mean_WP, data = final)
summary(WP_N)
```

```{r}
all %>%
  group_by(lab, Fish_pres) %>%
  summarize(total = n())
```

summarize the tanks
```{r}
metadata %>%
  filter(Fish_pres == 1) %>%
  filter(N_spp == 1) %>%
  filter(WP_pres == 1)
```


### explore residuals 

calculate and plot BIOMASS residuals
```{r}
residuals_biomass <- residuals(mod_all_biomass)
all_long_for_mod$residuals_biomass <- residuals_biomass

# Find indices of points with largest residuals
n_points <- 6  # Change this to the number of points you want to highlight
indices_of_largest_residuals <- order(abs(residuals_biomass), decreasing = TRUE)[1:n_points]

# Plot the scatter plot
plot(x = all_long_for_mod$Fbiomass, y = all_long_for_mod$Mean, main = "Scatter Plot with Line of Best Fit", xlab = "x", ylab = "y")
abline(mod_all_biomass, col = "red")  # Plot the line of best fit

# Highlight points with largest residuals
points(all_long_for_mod$Fbiomass[indices_of_largest_residuals], all_long_for_mod$Mean[indices_of_largest_residuals], col = "green", pch = 16)
```

```{r}
hist(residuals_biomass)

normparams <- MASS::fitdistr(residuals_biomass, "normal")$estimate                                      
probs <- pnorm(residuals_biomass, normparams[1], normparams[2])
outliers <- which(probs < 0.025 | probs > 0.975)

outlier_tanks <- all_long_abund_for_mod$tank_ID[outliers]
outlier_tanks
```

look at the details of the outlier tanks
```{r}
all_long %>%
  filter(tank_ID == 80)
```


since one of our objectives was to see how well the assay performed when species proportions were low vs middle vs high, let's look closer at that for the biomass model 

```{r}
biomass_resid_fig <- all_long_for_mod %>%
  mutate(Species = ifelse(Species == "AC", "Arctic cod", Species)) %>%
  mutate(Species = ifelse(Species == "PC", "Pacific cod", Species)) %>%
  mutate(Species = ifelse(Species == "WP", "Walleye pollock", Species)) %>%
  ggplot(aes(x = Fbiomass, y = residuals_biomass, color = Species)) +
  geom_point() +
  geom_smooth(aes(x=Fbiomass,y=residuals_biomass), color = "black", linetype = "dashed",method = "lm", se = TRUE) +
  scale_color_manual(values = mypal) + 
  geom_abline(intercept = 0, slope = 0, linetype = "dashed", color = "black")  + 
  theme_classic() +
  labs(x = "proportional biomass",
       y = "biomass model residuals") +
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14))

biomass_resid_fig
```     

** okay so a negative residual means there were more eDNA reads than true fish, and a positive residual means there was few reads than true fish 
** the most negative WP was a false positive detection in a take that should have been all AC.... 

```{r}
abs_biomass_resid_fig <- all_long_for_mod %>%
  mutate(Species = ifelse(Species == "AC", "Arctic cod", Species)) %>%
  mutate(Species = ifelse(Species == "PC", "Pacific cod", Species)) %>%
  mutate(Species = ifelse(Species == "WP", "Walleye pollock", Species)) %>%
  ggplot(aes(x = Fbiomass, y = abs(residuals_biomass), color = Species)) +
  geom_point() +
  geom_smooth(aes(x=Fbiomass,y= abs(residuals_biomass)), color = "black", linetype = "dashed",method = "lm", se = TRUE) +
  scale_color_manual(values = mypal) + 
  #geom_abline(intercept = 0, slope = 0, linetype = "dashed", color = "black")  + 
  theme_classic() +
  labs(x = "proportional biomass",
       y = "abs(biomass model residuals)") +
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14))

abs_biomass_resid_fig
```     



try to zoom in on the rare things
```{r}
fig_biomass_rare <- all_long %>%
  filter(n.tank.dissim != "2") %>%
  filter(Fish_pres == "1") %>%
  mutate(Species = ifelse(Species == "AC", "Arctic cod", Species)) %>%
  mutate(Species = ifelse(Species == "PC", "Pacific cod", Species)) %>%
  mutate(Species = ifelse(Species == "WP", "Walleye pollock", Species)) %>%
  filter(Fbiomass != 0 & Fbiomass <= 0.25) %>%
  ggplot() + 
  geom_errorbar(aes(x=Fbiomass,ymin=q.025,ymax=q.975,color=Species),width=0, size = .5) +
  geom_point(aes(x=Fbiomass,y=Mean,color=Species), size = 2) +
  #geom_smooth(aes(x=Fbiomass,y=Mean,color=Species), method = "lm", se = FALSE) +
  #geom_smooth(aes(x=Fbiomass,y=Mean), color = "black", linetype = "dashed", method = "lm", se = FALSE) +
  xlim(c(0, 0.6)) + 
  ylim(c(0, 0.6)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black")  + 
  scale_color_manual(values = mypal) + 
  theme_classic() + 
  labs(x = "proportional biomass",
       y = "eDNA read proportion") +
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14))
fig_biomass_rare
```


