---
title: "metabarcoding aquarium plots"
author: "Kimberly Ledger"
date: "2024-01-18"
output: html_document
---

this code is just a cleaned up version of the "7_mb_estimates_v_truth.Rmd" 

```{r}
library(tidyverse)
library(ggplot2)
select <- dplyr::select
```

read in tank estimates
```{r}
tanks_long <- read.csv("/home/kimberly.ledger/gadid_aquaria/gadid_mb_aquaria/tank_mock_estimates.csv") %>%
  mutate(Species = ifelse(Species == "Boreogadus saida", "AC", Species),
         Species = ifelse(Species == "Gadus chalcogrammus", "WP", Species),
         Species = ifelse(Species == "Gadus macrocephalus", "PC", Species))
```

read in tank metadata 
```{r}
metadata <-  read.csv("/home/kimberly.ledger/gadid_aquaria/gadid_mb_aquaria/TankData_MBmodified.csv") %>% 
  filter(groupID != "T") %>%  #remove tap water controls
  select(!sampleID) %>% #remove sample id because estimates are at the tank-level
  unique() %>%
  rename(tank_ID = groupID)
```

join - this removes the few T tanks with eDNA reads
```{r}
all_long <- metadata %>%
  pivot_longer(cols = c(19:21), names_to = "Species", values_to = "Fbiomass") %>%
  mutate(Species = str_remove(Species, "_Fbiomass")) %>%
  left_join(tanks_long, by = c("tank_ID", "Species"))
```

calculate abundance 
```{r}
all_long_abundance <- metadata %>%
  mutate(AC_N = ifelse(is.na(AC_N), 0, AC_N), 
         PC_N = ifelse(is.na(PC_N), 0, PC_N),
         WP_N = ifelse(is.na(WP_N), 0, WP_N),
         Tot_N = AC_N + PC_N + WP_N,
         AC_FN = AC_N/Tot_N,
         PC_FN = PC_N/Tot_N,
         WP_FN = WP_N/Tot_N) %>%
  pivot_longer(cols = c(33:35), names_to = "Species", values_to = "F_Abundance") %>%
  mutate(Species = str_remove(Species, "_FN")) %>%
  mutate(F_Abundance = ifelse(is.na(F_Abundance), 0, F_Abundance)) %>%
  left_join(tanks_long, by = c("tank_ID", "Species"))
```


plot biomass and color by species - only including samples with fish present 
```{r}
mypal <- c("#6a6599", "#79af97", "#00a1d5")   #purple, green, blue

fig_biomass <- all_long %>%
  filter(n.tank.dissim != "2") %>%
  filter(Fish_pres == "1") %>%
  mutate(Species = ifelse(Species == "AC", "Arctic cod", Species)) %>%
  mutate(Species = ifelse(Species == "PC", "Pacific cod", Species)) %>%
  mutate(Species = ifelse(Species == "WP", "Walleye pollock", Species)) %>%
  ggplot() + 
  geom_errorbar(aes(x=Fbiomass,ymin=q.025,ymax=q.975,color=Species),width=0, size = .5) +
  geom_point(aes(x=Fbiomass,y=Mean,color=Species), size = 2) +
  #geom_smooth(aes(x=Fbiomass,y=Mean,color=Species), method = "lm", se = FALSE) +
  geom_smooth(aes(x=Fbiomass,y=Mean), color = "black", linetype = "dashed", method = "lm", se = FALSE) +
  scale_color_manual(values = mypal) + 
  theme_classic() + 
  labs(x = "proportional biomass",
       y = "eDNA read proportion") +
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14))
fig_biomass
```

```{r}
ggsave("figures/read_biomass.png", plot = fig_biomass, width = 8.5, height = 6, dpi = 300)
```


plot abundance 
```{r}
fig_abund <- all_long_abundance %>%
  filter(n.tank.dissim != "2") %>%
  filter(Fish_pres == "1") %>%
  mutate(Species = ifelse(Species == "AC", "Arctic cod", Species)) %>%
  mutate(Species = ifelse(Species == "PC", "Pacific cod", Species)) %>%
  mutate(Species = ifelse(Species == "WP", "Walleye pollock", Species)) %>%
  ggplot() + 
  geom_errorbar(aes(x=F_Abundance,ymin=q.025,ymax=q.975,color=Species),width=0, size = 0.5) +
  geom_point(aes(x=F_Abundance,y=Mean,color=Species), size = 2) +
  #geom_smooth(aes(x=F_Abundance,y=Mean,color=Species,fill=Species), method = "lm", se = FALSE, alpha = 0.5) +
  geom_smooth(aes(x=F_Abundance,y=Mean), color = "black", linetype = "dashed",method = "lm", se = FALSE) +
  scale_color_manual(values = mypal) + 
  theme_classic() + 
  labs(x = "proportional abundance",
       y = "eDNA read proportion") +
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14))
fig_abund
```

```{r}
ggsave("figures/read_abund.png", plot = fig_abund, width = 8.5, height = 6, dpi = 300)
```


put together into one plot 
```{r, fig.width=10, fig.height=4}
library(patchwork)

combined_fig <- fig_biomass + fig_abund + plot_layout(guides = 'collect') + plot_annotation(tag_levels = 'A') &
    theme(plot.tag = element_text(face = 'bold'))
combined_fig
```

```{r}
ggsave("figures/biomass_abund.png", plot = combined_fig, width = 10, height = 4, dpi = 300)
```


linear regression with all species 
```{r}
all_long_for_mod <- all_long %>%
  filter(n.tank.dissim != "2") %>%
  filter(Fish_pres == "1")

mod_all_biomass <- lm(Fbiomass ~ Mean, data = all_long_for_mod)
summary(mod_all_biomass)

all_long_abund_for_mod <- all_long_abundance %>%
  filter(n.tank.dissim != "2") %>%
  filter(Fish_pres == "1")

mod_all_N <- lm(F_Abundance ~ Mean, data = all_long_abund_for_mod)
summary(mod_all_N)
```


make individual plots for each species on their own - using 1:1 line 
```{r}
fig_abund_PC_age <- all_long_abundance %>%
  filter(n.tank.dissim != "2") %>%
  filter(Fish_pres == "1") %>%
  mutate(Species = ifelse(Species == "AC", "Arctic cod", Species)) %>%
  mutate(Species = ifelse(Species == "PC", "Pacific cod", Species)) %>%
  mutate(Species = ifelse(Species == "WP", "Walleye pollock", Species)) %>%
  filter(Species == "Pacific cod") %>%
  ggplot() + 
  geom_errorbar(aes(x=F_Abundance,ymin=q.025,ymax=q.975,color=PC_age),width=0) +
  geom_point(aes(x=F_Abundance,y=Mean,color=PC_age), size = 3) + 
  #geom_errorbar(aes(x=F_Abundance,ymin=q.025,ymax=q.975,color=Species),width=0, size = 0.5) +
  #geom_point(aes(x=F_Abundance,y=Mean,color=Species), size = 3) +
  #geom_smooth(aes(x=F_Abundance,y=Mean,color=Species,fill=Species), method = "lm", se = FALSE, alpha = 0.5) +
  #geom_smooth(aes(x=F_Abundance,y=Mean), color = "black", linetype = "dashed",method = "lm", se = FALSE) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black")  + 
  scale_color_material("green") + 
  #scale_color_manual(values = "#79af97") +   #"#6a6599", "#79af97", "#00a1d5")   #purple, green, blue
  theme_classic() + 
  labs(x = "proportional abundance",
       y = "eDNA read proportion",
       color = "Pacific cod age") +
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14))
fig_abund_PC_age
```

```{r}
ggsave("figures/read_abund_PC_age.png", plot = fig_abund_PC_age, width = 8.5, height = 6, dpi = 300)
```


plot abundance - with biomass scale 
```{r}
temp <- all_long_abundance %>%
  filter(n.tank.dissim != "2") %>%
  filter(Fish_pres == "1") %>%
  mutate(AC_grams_per_N = AC_biomass/AC_N) %>%
  mutate(PC_grams_per_N = PC_biomass/PC_N) %>%
  mutate(WP_grams_per_N = WP_biomass/WP_N) %>%
  select(F_Abundance, Mean, q.025, q.975, Species, AC_grams_per_N, PC_grams_per_N, WP_grams_per_N) %>%
  filter(F_Abundance > 0) %>%
  pivot_longer(cols = c(6:8), values_to = "biomass_per_N") %>%
  mutate(name = ifelse(name == "AC_grams_per_N", "AC", name)) %>%
  mutate(name = ifelse(name == "PC_grams_per_N", "PC", name)) %>%
  mutate(name = ifelse(name == "WP_grams_per_N", "WP", name)) %>%
  filter(Species == name) %>%
  select(!name)
  
fig_abund_w_ind_biomass <- temp %>%  
  mutate(Species = ifelse(Species == "AC", "Arctic cod", Species)) %>%
  mutate(Species = ifelse(Species == "PC", "Pacific cod", Species)) %>%
  mutate(Species = ifelse(Species == "WP", "Walleye pollock", Species)) %>%
  filter(Species == "Pacific cod") %>%
  ggplot() + 
  geom_errorbar(aes(x=F_Abundance,ymin=q.025,ymax=q.975,color=Species),width=0, size = 0.5) +
  geom_point(aes(x=F_Abundance,y=Mean,color=Species, fill=biomass_per_N), shape = 21, size = 2, stroke = 1) +
  #geom_smooth(aes(x=F_Abundance,y=Mean), color = "black", linetype = "dashed",method = "lm", se = FALSE) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black")  + 
  scale_color_manual(values = "#79af97") +   #"#6a6599", "#79af97", "#00a1d5"   #purple, green, blue
  #scale_color_manual(values = mypal) + 
  #scale_fill_continuous(low = "white", high = "black", trans = "log10") +  # Apply log transformation to fill scale
  scale_fill_continuous(low = "white", high = "black", trans = "sqrt") +  # Apply log transformation to fill scale
  theme_classic() + 
  labs(x = "proportional abundance",
       y = "eDNA read proportion",
       fill = "grams per individual") +
  #theme(axis.text = element_text(size = 12),
  #      axis.title = element_text(size = 14),
  #      legend.text = element_text(size = 12),
  #      legend.title = element_text(size = 14)) + 
  guides(color = "none")
fig_abund_w_ind_biomass
```

```{r}
ggsave("figures/read_abund_PCOD_biomass.png", plot = fig_abund_w_ind_biomass, width = 8.5, height = 6, dpi = 300)
```



```{r}
temp2 <- all_long %>%
  filter(n.tank.dissim != "2") %>%
  filter(Fish_pres == "1") %>%
  mutate(AC_grams_per_N = AC_biomass/AC_N) %>%
  mutate(PC_grams_per_N = PC_biomass/PC_N) %>%
  mutate(WP_grams_per_N = WP_biomass/WP_N) %>%
  select(Fbiomass, Mean, q.025, q.975, Species, AC_grams_per_N, PC_grams_per_N, WP_grams_per_N) %>%
  filter(Fbiomass > 0) %>%
  pivot_longer(cols = c(6:8), values_to = "biomass_per_N") %>%
  mutate(name = ifelse(name == "AC_grams_per_N", "AC", name)) %>%
  mutate(name = ifelse(name == "PC_grams_per_N", "PC", name)) %>%
  mutate(name = ifelse(name == "WP_grams_per_N", "WP", name)) %>%
  filter(Species == name) %>%
  select(!name)
  
fig_biomass_w_ind_biomass <- temp2 %>%  
  mutate(Species = ifelse(Species == "AC", "Arctic cod", Species)) %>%
  mutate(Species = ifelse(Species == "PC", "Pacific cod", Species)) %>%
  mutate(Species = ifelse(Species == "WP", "Walleye pollock", Species)) %>%
  filter(Species == "Pacific cod") %>%
  ggplot() + 
  geom_errorbar(aes(x=Fbiomass, ymin=q.025,ymax=q.975,color=Species),width=0, size = 0.5) +
  geom_point(aes(x=Fbiomass,y=Mean,color=Species, fill=biomass_per_N), shape = 21, size = 2, stroke = 1) +
  #geom_smooth(aes(x=F_Abundance,y=Mean), color = "black", linetype = "dashed",method = "lm", se = FALSE) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black")  + 
  scale_color_manual(values = "#79af97") +   #"#6a6599", "#79af97", "#00a1d5"   #purple, green, blue
  #scale_color_manual(values = mypal) + 
  #scale_fill_continuous(low = "white", high = "black", trans = "log10") +  # Apply log transformation to fill scale
  scale_fill_continuous(low = "white", high = "black", trans = "sqrt") +  # Apply log transformation to fill scale
  theme_classic() + 
  labs(x = "proportional biomass",
       y = "eDNA read proportion",
       fill = "grams per individual") +
  #theme(axis.text = element_text(size = 12),
  #      axis.title = element_text(size = 14),
  #      legend.text = element_text(size = 12),
  #      legend.title = element_text(size = 14)) + 
  guides(color = "none")
fig_biomass_w_ind_biomass
```




now explore the relationship between mass and the read proportion v abundance proportion residuals 

having to run code from "10_estimtes_v_truth" scripts because i don't know why i cant get things to work...

read in tank estimates
```{r}
tanks <- read.csv("/home/kimberly.ledger/gadid_metabarcoding/aquaria_DBO/tank_mock_estimates.csv") %>%
  mutate(Species = ifelse(Species == "Boreogadus saida", "AC", Species),
         Species = ifelse(Species == "Gadus chalcogrammus", "WP", Species),
         Species = ifelse(Species == "Gadus macrocephalus", "PC", Species)) %>%
  pivot_wider(names_from = Species, values_from = c(6:9))
```

read in tank metadata 
```{r}
metadata <-  read.csv("/home/kimberly.ledger/gadid_metabarcoding/aquaria_DBO/TankData_MBmodified.csv") %>% 
  filter(groupID != "T") %>%  #remove tap water controls
  select(!sampleID) %>% #remove sample id because estimates are at the tank-level
  unique() %>%
  rename(tank_ID = groupID)
```

join - this removes the few T tanks with eDNA reads. okay for now but will need to report those somewhere... 
```{r}
all <- metadata %>%
  left_join(tanks, by = "tank_ID")
```

```{r}
all$n.tank.dissim <- as.factor(all$n.tank.dissim)
all$n.sample.dissim <- as.factor(all$n.sample.dissim)
```

```{r}
all <- all %>%
  mutate(AC_N = ifelse(is.na(AC_N), 0, AC_N), 
         PC_N = ifelse(is.na(PC_N), 0, PC_N),
         WP_N = ifelse(is.na(WP_N), 0, WP_N),
         Tot_N = AC_N + PC_N + WP_N,
         AC_FN = AC_N/Tot_N,
         PC_FN = PC_N/Tot_N,
         WP_FN = WP_N/Tot_N)
```

compute the perpendicular distance of each point to the line
```{r}
dist_data <- all %>%
  filter(n.tank.dissim != "2") %>%
  filter(Fish_pres == "1") %>%
  mutate(AC_grams_per_N = AC_biomass/AC_N) %>%
  mutate(PC_grams_per_N = PC_biomass/PC_N) %>%
  mutate(WP_grams_per_N = WP_biomass/WP_N)

# Calculate distances from points to the 1:1 line
dist_data$PC_distances_abundance <- dist_data$Mean_PC - dist_data$PC_FN
dist_data$PC_distances_biomass <- dist_data$Mean_PC - dist_data$PC_Fbiomass
dist_data$AC_distances_abundance <- dist_data$Mean_AC - dist_data$AC_FN
dist_data$AC_distances_biomass <- dist_data$Mean_AC - dist_data$AC_Fbiomass
dist_data$WP_distances_abundance <- dist_data$Mean_WP - dist_data$WP_FN
dist_data$WP_distances_biomass <- dist_data$Mean_WP - dist_data$WP_Fbiomass

#dist_data <- dist_data %>%
#  filter(!is.na(PC_age))
dist_data
```


compute the perpendicular distance of each point to the line
```{r}
# Calculate distances from points to the 1:1 line
dist_data$distances_abundance <- dist_data$Mean_PC - dist_data$PC_FN
dist_data$distances_biomass <- dist_data$Mean_PC - dist_data$PC_Fbiomass

dist_data <- dist_data %>%
  filter(!is.na(PC_age))
```


```{r}
dist_data <- dist_data %>%
  mutate(PC_age_group = ifelse(PC_age == 0, "age 0", NA),
         PC_age_group = ifelse(PC_age == 1, "age 1", PC_age_group),
         PC_age_group = ifelse(PC_age == 1.5, "age 1", PC_age_group),
         PC_age_group = ifelse(PC_age == 2, "age 2", PC_age_group),
         PC_age_group = ifelse(PC_age == 7, "age 7", PC_age_group))
```


```{r}
pc_residual_abund <- ggplot(dist_data, aes(y = PC_age_group, x = distances_abundance, fill = PC_age)) + 
  geom_point(shape = 21, cex = 2) + 
  theme_bw() + 
  scale_fill_material("green") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  labs(x = "residual (abundance)",
       y = "Pacific cod age") +
  theme(legend.position = "none")+
  xlim(c(-0.5,0.5))
pc_residual_abund
```

```{r}
pc_residual_biomass<- ggplot(dist_data, aes(y = PC_age_group, x = distances_biomass, fill = PC_age)) + 
  geom_point(shape = 21, cex = 2) + 
  theme_bw() + 
  scale_fill_material("green") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  labs(x = "residual (biomass)",
       y = "Pacific cod age") +
  theme(legend.position = "none")+
  xlim(c(-0.5,0.5))
pc_residual_biomass
```



now plot mass instead of age 

```{r}
pc_residual_abund_by_mass <- ggplot(dist_data, aes(y = PC_grams_per_N, x = distances_abundance, fill = PC_age)) + 
  geom_point(shape = 21, cex = 2) + 
  theme_bw() + 
  scale_fill_material("green") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  scale_y_sqrt() +
  labs(x = "residual (abundance)",
       y = "Pacific cod mass (g)",
       fill = "Pacific cod age") +
  theme(legend.position = "right")+
  xlim(c(-0.5,0.5))
pc_residual_abund_by_mass
```

```{r}
pc_residual_biomass_by_mass <- ggplot(dist_data, aes(y = PC_grams_per_N, x = distances_biomass, fill = PC_age)) + 
  geom_point(shape = 21, cex = 2) + 
  theme_bw() + 
  scale_fill_material("green") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  scale_y_sqrt() +
  labs(x = "residual (biomass)",
       y = "Pacific cod mass (g)",
       fill = "Pacific cod age") +
  theme(legend.position = "right")+
  xlim(c(-0.5,0.5))
pc_residual_biomass_by_mass
```

combined figure 
```{r}
prop_fig <- fig_biomass_w_ind_biomass + fig_abund_w_ind_biomass + plot_layout(guides = 'collect') 
prop_fig

residual_fig <-  pc_residual_biomass_by_mass+ pc_residual_abund_by_mass + plot_layout(guides = 'collect') 
residual_fig
```

```{r}
all_together <- prop_fig / residual_fig + plot_annotation(tag_levels = 'A') &
    theme(plot.tag = element_text(face = 'bold'))
all_together
```

```{r}
#ggsave("figures/Pcod_biomass_abund_w_residuals.png", plot = all_together, width = 8.5, height = 6, dpi = 300)
```

